Project Notes

### 2020-04-10 notes/notes.tex
Notes: start define formalism for Node send/receive

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}

@@

\item
Could also look up source code of Scribble-Java to see what they use.
\end{itemize}

@@

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath}

@@

\item
Could also look up source code of Scribble-Java to see what they use.

\item
Fix message payload parsing in EFSM to correctly strip and filter out whitespace.

\item
Adapt Node/React templates for message payload types -- the Calculator example
(binary session with recursion, selection and branching) now produces valid TypeScript
code for both browser and client.

@@

\subsubsection{Encoding}
We distinguish between \textbf{handler} (the PLACES encoding which handle the
message sending and receiving) and \textbf{implementation} (the wrapper around the handler).

\paragraph{Send / Selection}
Assume state S1 can select from a list of [Label, Payload, Successor] such that Payload and Successor
are dependent on Label.

\begin{align}
encoding_{Send}(S) &= handler_{Send}(S), impl_{Send}(S) \\
handler_{Send}(S) &= \bigcup_{i \in labels(S)} [Labels_i,~Payload_i,~Impl(Succ_i)] \\
impl_{Send}(S) &= sendImpl(Handler(S),~Message(S))
\end{align}

\texttt{sendImpl} defines a subclass of \texttt{Implementation.Send} based on the handler data type
and message data type (the union of possible messages). It implements the abstract method
\texttt{performSend} based on these parameters, which (1) sends the message using
the method (or channel passed in by runtime), then (2) advances to next state (again, using
function passed in by runtime which acts as a ``send``).

\paragraph{Receive / Branching}

\begin{align}
encoding_{Recv}(S) &= handler_{Recv}(S), impl_{Recv}(S) \\
handler_{Recv}(S) &= Map_{i \in labels(S)}\left<Labels_i, Function\left<Payload_i, Impl(Succ_i)\right>\right> \\
impl_{Recv}(S) &= recvImpl(Handler(S), Message(S))
\end{align}

\texttt{recvImpl} defines a subclass of \texttt{Implementation.Recv} based on the handler data type
and message data type (the union of possible messages). It implements the abstract method
\texttt{performRecv} based on these parameters, which (1) prepares the receive handler (a function that takes
(message: any) => void), then (2) registers this handler to the runtime using the passed-in callback
(which acts as sending the handler to the runtime), and (3) when said handler in (1) gets called by the runtime
(essentially receiving from the runtime), actually performs the receive and (4) advances to next state.

\subsubsection{Runtime as session type}
Loosely define the protocol for message passing between the runtime and the
corresponding state implementation.

\begin{lstlisting}[breaklines=true, tabsize=4, frame=single]
protocol Next(role Runtime, role State) {
	// not really a `choice`, more of `based on the type of State`\dots
	choice at Runtime {
		// non-terminal
		choice at Runtime {
			// send
			PerformSend(???) from Runtime to State;
			Send(role: string, label: string, payload: any[]) from State to Runtime;

			// runtime sends message through WebSocket

			// `Successor` is the state transition function in the literature
			do Next(Runtime, Successor<State, label>);

		} or {
			// receive
			PrepareReceive(???) from Runtime to State;
			do Receive(Runtime, State);
		}

	} or {
		// terminal
	}
}

protocol Receive(role Runtime, role State) {

	/*
	choice {
		Message(label: string, payload: any[]) from ??? to Runtime;
		Register(???) from State to Runtime;
		PerfomReceive(???) from Runtime to State;
		do Next(Runtime, Successor<State, label>);
	} or {
		Register(???) from State to Runtime;
		Message(label: string, payload: any[]) from ??? to Runtime;
		PerformReceive(???) from Runtime to State;
		do Next(Runtime, Successor<State, label>); 
	}
	*/

	// State sends a Channel to Runtime
	Register(performReceive: ???) from State to Runtime;

	// whenever the message arrives -- regardless whether it is before or after
	performReceive({ label: string, payload: any[] }) from Runtime to State;

	do Next(Runtime, Successor<State, label>);
}
\end{lstlisting}
